Fitting Synchrotron Models to Gamma-Ray Burst Data from INTEGRAL/SPI

### 2021-12-16 src/ms.tex
test citation

\label{firstpage}
% Abstract with filler text
\abstract {
  In the last years it has been shown, that it is needed to fit physical models to Gamma-Ray Burst (GRB) data to get more insights into their emission mechanism. For the energy range between ~ 10 keV to ~ 10 MeV Fermi/GBM has been the main instrument.... In this paper we show how we can add INTEGRAL/SPI to this analysis. INTEGRAL/SPI covers a similar energy range like Fermi/GBM and has a very good energy resolution of 2.2 keV at 1.3 MeV. Therefore INTEGRAL/SPI is an ideal instrument to precisely measure the curvature of the spectrum. In this work we present PySPI, a new analysis software we developed to analyze GRB data from SPI, and apply it to the GRB 120711A. We show that PySPI improves the analysis of SPI data, for GRBs, compared to the official analysis software. Also we show that the GBM and the SPI data of this GRB can be fitted well with a physical synchrotron model and that combining both, reduces the allowed parameter space.
}

@@

Gamma-Ray Bursts (GRBs) are short transient bursts of gamma rays, with a typical active time range between a few ms and few hundred seconds for the prompt phase, which is mostly observed in X- and Gamma-Rays. After the prompt emission there is a long lasting afterglow phase when the ejecta interacts with the Inter Stellar Medium (ISM). Long GRBs ($\approx$> 2s) have been associated to the collapse of massive stars (cite), whereas short GRBs are believed to be caused by mergers of compact objects, like Neutron Stars (cite grb-gw). In both cases it is commonly agreed on, that the progenitor events of GRBs produce jetted relativistic outflows, which consists of several shells with different velocities, resulting in internal shell collisions. But it is still highly debated what the exact emission mechanism of the Gamma-Rays is.

In the past, GRB data was often fitted with empirical function like the Band function (cite band). But in recent years it has been shown, that this approach can be misleading and that one should rather fit physical models directly to the data. An example for this is the so-called line-of-death for synchrotron emission, which states that synchrotron radiation can not be the emission mechanism, if fits with a Band function give a low energy power law slope larger than -2/3. (cite) have shown that this conclusion is not correct and that you can fit GRB spectra well with a physical synchrotron model even though the same spectra violate the line-of-death if fitted with a Band function.
The Gamma-Ray Burst Monitor (GBM) onboard of the Fermi satellite is often used to fit GRB data in the Gamma-Ray energy range. GBM is an all-sky instrument, that observes the whole sky, except the part that is occulted by the Earth at that time. It consists of 12 NaI and two BGO scintillator detectors that cover in total an energy range between 8 keV and 10 MeV. (cite)
%The problem with fitting empirical models to the data and using the individual parameters of these fits to construct constrains on the physics is that the
Another instrument that operates in a similar energy range is the SPectrometer on INTEGRAL (SPI), which is one of the instruments onboard of the INTErnational Gamma-Ray Astrophysics Laboratory (INTEGRAL) satellite, that was launched in 2002 and is still operating today. SPI is a coded mask instrument, which has a FOV of roughly ? sr and covers an energy range between 20 keV and 8 MeV. Due to the Germanium detectors it has a very good energy resolution of roughly 2.2 keV at 1.3 MeV, which allows the identification of fine features, like atomic decay lines. (cite) In the case of GRBs this excellent energy resolution could allow us to determine the curvature of the spectra with an unmet precision. This is the main motivation to use SPI in our analysis.

@@

GRBs are transient sources with a typical duration only up to a few tens of seconds.
%Therefore we can use the time before and after the GRB signal, but within the same science window, as a background measurement.
%The background treatment we use, depends on two assumptions, namely that within one science window the background is stable and the pointing direction of SPI does not change.
%Due to the high orbit and pointing scheme of INTEGRAL both these assumptions are valid.
Therefore we can use the time intervals during a science window when the transient source is not active as an independent temporal off-source observation (see Fig. \ref{fig:lightcurve}). Similar to other instruments that use spatially off-source observation as an independent background measurement. From the background observations we construct the probability of the background model rates per energy channel from a poisson distribution, like given in equation \ref{eq:poisson_bkg}, where $b_{i}$ are the background rates per energy channel, $B_{i}$ are the detected counts in the off-source observation and $t_{b}$ is the exposure of the off-source observation.
\begin{equation}
	L(b_{i}|B_{i}, t_{b})=\frac{(t_{b} b_{i})^{B_{i}}}{B_{i}!}\cdot e^{-t_{b} b_{i}}
  \label{eq:poisson_bkg}
\end{equation}
This expression will be used in the next paragraph to derive the total likelihood.

@@

Because there is no easy analytical solution for this integral, one often uses a so-called profiled likelihood as an approximation. For the profiled likelihood we use the fact, that the derivative of the likelihood at its maximum is zero. Therefore one can construct from $\frac{\textrm{dL}}{\textrm{db}_i}=0$ the background rates $b_{i, max}$ that maximize the likelihood for a given model $m_{i}(\theta)$ and observed counts. Solving $\frac{\textrm{dL}}{\textrm{db}_i}=0$ for b gives equation \ref{eq:bmax}.
\begin{equation}
	b_{i,max}(\theta)=\frac{1}{2(t_{b}+t_{d})}(B_{i}+D_{i}-m_{i}(\theta)(t_{b}+t_{d})+\sqrt{(B_{i}+D_{i}+m_{i}(\theta)(t_{b}+t_{d}))^{2}-4m_{i}(\theta)D_{i}(t_{d}-t_{b})}
  \label{eq:bmax}
\end{equation}

These values can then be substituted in the likelihood \ref{eq:likelihood_m_and_b} to eliminate the $b_{i}$ dependency. This approximation gives the exact solution for the maximum of the likelihood, for likelihood values close to the maximum the assumption is that most of the likelihood in the integrand in \ref{eq:likelihood_marg} is in a small area around $b_{i,max}$ and we can basically add a $\delta(b_{i}-b_{{i,max}})$. This assumption is used in many works (cite) and works very well as long as one does not have too few counts in the background observation (for a derivation see cite Giacomo derivation).

@@

These effective areas were calculated for 51 photon energies and on a 0.5 degree grid. For each of the three interaction type there is one RMF, to define the shape of the redistributed spectra, assuming that this shape does not depend on the detector or the incident angles of the photons.
The procedure to construct the response includes several simplification to keep the computational costs and storage space manageable (cite). This could of course affect every analysis done with SPI. Re-simulating the response without these simplifications could improve the scientific output of SPI, but would take a hugh amount of computational time, even today.

@@

Since shortly after the start of INTEGRAL, it is known that there are spurious events in the SPI data around 1.5 MeV. According to (cite Roques) these spurious events are photons with small energy (<100 keV) that get detected at a higher energy due to saturation of the ??? electronics by previous high energy deposition. It is also known, that these spurious events do not show up if one looks only at events with a detection also in the Pulse Shape Discriminator (PSD) electronics. The reason for this is that the PSD electronics has an low energy threshold of roughly 450 keV (the exact value has been changed a few times during the mission). Therefore only events that deposit more than this low energy threshold in the Germanium crystal can trigger this electronic, which eliminates the <100 keV events that are detected at the wrong energies by the AFEE electronics. In Fig. \ref{fig:electronic} one can see that the feature at 1.5 MeV is nicely visible in the non-PSD events but is missing in the PSD events. Even though this problem is the most significant in the area around 1.5 MeV it is also important at lower energies >400 keV. In PySPI one can select the energy range in which only the PSD events should be used, to avoid including spurious events. To account for the larger dead time of the PSD electronics, an effective area correction can be either fix to 85 \% (cite) or treated as a free parameter in the fit.

@@

In PySPI every Germanium detector is treated as an independent detector. The workflow during a fit step is the following:
\begin{itemize}
  \item Sample model parameters
  \item Calculate model flux and responses individually for all detectors for the source position
  \item Fold model with responses to get predicted model counts in all detectors
  \item Calculate log-likelihood for all detectors
  \item Sum these log-likelihoods to get the total log-likelihood of SPI
\end{itemize}
A sampling algorithm like emcee (cite) or MultiNest (cite) is used to iterate this procedure, in order to determine the posterior distributions.\\
This workflow is very similar, like for example for GBM, because fundamentally instruments like SPI and GBM are very similar. They both consist of individual detectors that have different responses for a given source position. In GBM this is due to the different pointing directions of the detectors and in SPI due to the different occultation by the mask. Therefore there is no reason to treat SPI specially, just because it is a coded mask instrument. The information the mask encodes into the data is automatically included by using the different responses for the detectors.

@@

PySPI can construct a plugin for the Multi-Mission Maximum Likelihood framework (3ML). This allows, without any further work, to fit the SPI data together with the data from other instruments, like Fermi/GBM. For more information about 3ML see (cite paper and documentation).

@@

PySPI is an open source software, which is publicly available on GitHub. There is also a documentation, that shows how to use it. (cite JOSS, link to repo, docs).

@@

As we are introducing a new analysis software in this work, we want to compare it to the existing analysis software, to check if the results are in agreement. We shortly summarize the official analysis tools for GRB analysis, which are part of the INTEGRAL Offline Scientific Analysis (OSA) software. (https://ui.adsabs.harvard.edu/abs/2003A%26A...411L..53C/abstract)

@@

The pseudo photon flux data points from the 'Photo Peak Fitting' step, which are no real data points anymore, are fitted with a spectral photon flux model which is convolved with a correction response, which is also available through OSA. This correction response depends on the source position and energy bins that are used and should account for energy dispersion as well as resolution. In this step the energy channels are fitted simultaneously, as it is impossible to incorporate energy dispersion in a way that fits every energy channel individually. This fitting can be done in different spectral fitting software like 3ML or XSPEC. In this work we use 3ML.

\subsection{Synchrotron Model}
\label{synch}
We use pynchrotron (cite github) to calculate the spectrum from a physical, time dependent spectral synchrotron model. This model was previously used to successfully fit many single pulse GRBs with GBM in cite syn paper. This paper also include a detailed description of the model. Here we only want to summarize the main points shortly.

The core of the model is the assumption of some generic acceleration mechanism, that constantly produces an electron spectrum with a power law shape $N(\gamma )\propto gamma^{-p}$ between $\gamma_{inj}$ and $\gamma_{max}$. These electrons are cooled via synchrotron cooling in a magnetic field and produce radiation in this process. The resulting photon spectrum is the sum of the photon spectra at every time step in the cooling process and is defined by five quantities:

@@

There exists a strong degeneracy between $B$ and $\gamma_{inj}$, as their combination sets the peak of the photon spectrum. Therefore one usually fixes $\gamma_{inj}=10^{5}$ and only fits for $B$. It is important to keep in mind, that the results for $\gamma_{max}$ and $\gamma_{cool}$ are determined relative to the fixed $\gamma_{inj}$. For more details about the model check (cite Syn paper).

\subsection{Model Checking}
\label{PPC}
To check if the fit of a model is a good description of the measured data is a very complicated topic and highly debated in the statistics community. Measurements like reduced $\chi^{2}$ use many assumptions, like for example that all the probability distributions in the problem are described as normal distributions, which is definitely not the case here, as the measurement process is a Poisson process. Therefore we decided to use Posterior Predictive Checks (PPC) and quantile-quantile (QQ) plots. For the PPCs one simulate new data from the full posterior of the fit as well as the measurement process (see Eq. \ref{eq:ppc}) and compare them to the observed data. QQ plots use the same simualtion process, but instead of comparing the observed data of every energy channel individually to the simulated data, one compares the cumulative counts of the observation and the simulations over energy channels. QQ plots are very sensitive to weak systematic deviations of the model from the data.

@@

We analyze the data from the two instruments independently first with the empirical Band function, to check if the results for SPI and GBM are in agreement as well as show a comparison between the fit results with PySPI and the official software tools for GRB analysis within OSA. Then we fit the SPI and GBM data simultaneously with a physical synchrotron model, and show that it can reduce the allowed parameter space significantly, compared to the individual fits.

All the fits in this section are done with 3ML (cite) and the Bayesian sampler algorithm MultiNest (cite) to create full posterior distributions of the parameters. To visualize the posterior distribution we use ChainConsumer (cite) with the output from MultiNest to create corner plots.

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=1\linewidth]{figures/pyspi_vs_osa_band.pdf}
    \caption{Corner plot for the SPI fit with PySPI and the official tools. The spectral model in this fit is a band function. It shows, that the results are in agreement, but PySPI can constrain the parameters better.}
    \label{fig:corner_osa_pyspi_band}
  \end{centering}
\end{figure}

@@

We analyzed the data of GRB120711A with the data from INTGRAL/SPI and Fermi/GBM. The results for SPI with PySPI are in agreement with those from the official analysis software, but more constraining. Also the resulting spectral shape for the fits with GBM and SPI data are in agreement and the relative calibration difference between SPI and GBM we conclude from the fits is $\approx$ 10\% or less. We show that the GBM as well as the SPI data for GRB120711A can be fitted well with a physical, time dependent synchrotron model and that combining the data in a simultaneous fit improve the parameter constrains.

Generally PySPI can improves the scientific output from SPI for GRBs significantly. Due to a new analysis method, that folds the source model directly into the data space of every SPI detector with the full response of the individual detectors, the results are more constraining than the results with the official analysis software.

>>>>

\label{firstpage}
% Abstract with filler text
\abstract {
  In the last years it has been shown, that it is needed to fit physical models to Gamma-Ray Burst (GRB) data to get more insights into their emission mechanism. For the energy range between ~ 10 keV to ~ 10 MeV Fermi/GBM has been the main instrument to analyze GRBs. In this paper we show how we can add INTEGRAL/SPI to this analysis. INTEGRAL/SPI covers a similar energy range like Fermi/GBM and has a very good energy resolution of 2.2 keV at 1.3 MeV. Therefore INTEGRAL/SPI is an ideal instrument to precisely measure the curvature of the spectrum. We present PySPI, a new analysis software we developed to analyze GRB data from SPI, and apply it to the GRB 120711A. We show that PySPI improves the analysis of SPI data, for GRBs, compared to the official analysis software and that the GBM and the SPI data of this GRB can be fitted well with a physical synchrotron model and that combining both, reduces the allowed parameter space. This demonstrates that SPI can play an important role in GRB model fitting.
}

@@

Gamma-Ray Bursts (GRBs) are short transient bursts of gamma rays, with a typical active time range between a few ms and few hundred seconds for the prompt phase, which is mostly observed in X- and Gamma-Rays. After the prompt emission, there is a long lasting afterglow phase when the ejecta interacts with the Inter Stellar Medium (ISM). Long GRBs ($\approx$> 2s) have been associated to the collapse of massive stars, whereas short GRBs are believed to be caused by mergers of compact objects, like Neutron Stars. In both cases it is commonly agreed on, that the progenitor events of GRBs produce jetted relativistic outflows, which consists of several shells with different velocities, resulting in internal shell collisions. But it is still highly debated what the exact emission mechanism of the Gamma-Rays is. \citep{gamma-ray-burst, gw}

In the past, GRB data was often fitted with empirical function like the Band function (cite band). But in recent years it has been shown, that this approach can be misleading and that one should rather fit physical models directly to the data. An example for this is the so-called line-of-death for synchrotron emission, which states that synchrotron radiation can not be the emission mechanism, if fits with a Band function give a low energy power law slope larger than -2/3. (cite) have shown that this conclusion is not correct and that you can fit GRB spectra well with a physical synchrotron model even though the same spectra violate the line-of-death if fitted with a Band function.

The Gamma-Ray Burst Monitor (GBM) onboard of the Fermi satellite is often used to fit GRB data in the 8 keV to 10 MeV energy range. GBM consists of 12 NaI and two BGO scintillator detectors and is an all-sky instrument, that observes the whole sky, except the part that is occulted by the Earth at that time. (cite)
%The problem with fitting empirical models to the data and using the individual parameters of these fits to construct constrains on the physics is that the
Another instrument that operates in a similar energy range is the SPectrometer on INTEGRAL (SPI), which is one of the instruments onboard of the INTErnational Gamma-Ray Astrophysics Laboratory (INTEGRAL) satellite, that was launched in 2002 and is still operating today. SPI is a coded mask instrument, which has a FOV of roughly ? sr and covers an energy range between 20 keV and 8 MeV. Due to the Germanium detectors it has a very good energy resolution of roughly 2.2 keV at 1.3 MeV, which allows the identification of fine features, like atomic decay lines. (cite) In the case of GRBs this excellent energy resolution could allow us to determine the curvature of the spectra with an unmet precision. This is the main motivation to use SPI in our analysis.

@@

GRBs are transient sources with a typical duration only up to a few tens of seconds. Therefore we can use the time intervals during a science window, when the transient source is not active, as an independent temporal off-source observation (see Fig. \ref{fig:lightcurve}). Similar to other instruments, that use spatially off-source observation as an independent background measurement. From the background observations we construct the probability of the background model rates per energy channel from a poisson distribution, like given in equation \ref{eq:poisson_bkg}, where $b_{i}$ are the background rates per energy channel, $B_{i}$ are the detected counts in the off-source observation and $t_{b}$ is the exposure of the off-source observation.
\begin{equation}
	L(b_{i}|B_{i}, t_{b})=\frac{(t_{b} b_{i})^{B_{i}}}{B_{i}!}\cdot e^{-t_{b} b_{i}}
  \label{eq:poisson_bkg}
\end{equation}
This expression will be used in the next paragraph to derive the total likelihood.

@@

Because there is no easy analytical solution for this integral, one often uses a so-called profiled likelihood as an approximation (cite). For the profiled likelihood we use the fact, that the derivative of the likelihood at its maximum is zero. Therefore one can construct from $\frac{\textrm{dL}}{\textrm{db}_i}=0$ the background rates $b_{i, max}$ that maximize the likelihood for a given model $m_{i}(\theta)$ and observed quantities. Solving $\frac{\textrm{dL}}{\textrm{db}_i}=0$ for b gives equation \ref{eq:bmax}.
\begin{equation}
	b_{i,max}(\theta)=\frac{1}{2(t_{b}+t_{d})}(B_{i}+D_{i}-m_{i}(\theta)(t_{b}+t_{d})+\sqrt{(B_{i}+D_{i}+m_{i}(\theta)(t_{b}+t_{d}))^{2}-4m_{i}(\theta)D_{i}(t_{d}-t_{b})}
  \label{eq:bmax}
\end{equation}

These values can then be substituted in the likelihood \ref{eq:likelihood_m_and_b} to eliminate the $b_{i}$ dependency. This approximation gives the exact solution for the maximum of the likelihood, for likelihood values close to the maximum, the assumption is that most of the likelihood in the integrand in \ref{eq:likelihood_marg} is in a small area around $b_{i,max}$. This assumption is used in many spectral analyses and works very well as long as one does not have too few counts in the background observation \footnote{For a derivation see https://giacomov.github.io/Bias-in-profile-poisson-likelihood/}.

@@

These effective areas were simulated for 51 photon energies and on a 0.5 degree grid. For each of the three interaction type, there is one RMF, to define the shape of the redistributed spectra, assuming that this shape does not depend on the detector or the incident angles of the photons. The procedure to construct the response includes several simplification to keep the computational costs and storage space manageable (cite). This could of course affect every analysis done with SPI. Re-simulating the response without these simplifications could improve the scientific output of SPI, but would take a hugh amount of computational time, even today.

@@

Since shortly after the start of INTEGRAL, it is known that there are spurious events in the SPI data around 1.5 MeV. According to (cite Roques) these spurious events are photons with small energy (<100 keV) that get detected at a higher energy due to saturation of the Analog Front-End Electronics (AFEE) electronics by previous high energy deposition. It is also known, that these spurious events do not show up if one looks only at events with a detection also in the Pulse Shape Discriminator (PSD) electronics. The reason for this is that the PSD electronics has a low energy threshold of roughly 450 keV (the exact value has been changed a few times during the mission). Therefore only events that deposit more than this low energy threshold in the Germanium crystal can trigger this electronic, which eliminates the <100 keV events that are detected at the wrong energies by the AFEE electronics. In Fig. \ref{fig:electronic} one can see that the feature at 1.5 MeV is nicely visible in the non-PSD events but is missing in the PSD events. Even though this problem is the most significant in the area around 1.5 MeV it is also important at lower energies >400 keV. In PySPI one can select the energy range in which only the PSD events should be used, to avoid including spurious events. To account for the larger dead time of the PSD electronics, an effective area correction can be either fix to 85 \% (cite) or treated as a free parameter in the fit.

@@

In PySPI every Germanium detector is treated as an independent detector. The workflow during a fit step is the following:
\begin{itemize}
  \item Sample model parameters
  \item Calculate model flux and responses individually for all detectors for the source position
  \item Fold model with responses to get predicted model counts in all detectors
  \item Calculate log-likelihood for all detectors
  \item Sum these log-likelihoods to get the total log-likelihood of the SPI data for the given model parameters
\end{itemize}
A sampling algorithm like MultiNest (cite) is used to iterate this procedure, in order to determine the posterior distributions.\\
This workflow is very similar, like for example for GBM, because fundamentally instruments like SPI and GBM are very similar. They both consist of individual detectors that have different responses for a given source position. In GBM this is due to the different pointing directions of the detectors and in SPI due to the different occultation by the mask. Therefore there is no reason to treat SPI specially, just because it is a coded mask instrument. The information the mask encodes into the data is automatically included by using the different responses for the detectors.

@@

PySPI can construct a plugin for the Multi-Mission Maximum Likelihood framework (3ML). This allows, without any further work, to fit the SPI data together with the data from other instruments, like Fermi/GBM. \footnote{For more information about 3ML see threeml.readthedocs.io/}

@@

PySPI is an open source software, which is publicly available on GitHub (cite JOSS). There is also a documentation, that shows how to use it. \footnote{For more information about PySPI see pyspi.readthedocs.io/}

@@

As we are introducing a new analysis software in this work, we want to compare it to the existing analysis software, to check if the results are in agreement. We shortly summarize the official analysis tools for GRB analysis, which are part of the INTEGRAL Offline Scientific Analysis (OSA) software. (cite osa)

@@

The pseudo photon flux data points from the 'Photo Peak Fitting' step, which are no real data points anymore, are fitted with a spectral photon flux model which is convolved with a correction response, which is also available through OSA. This correction response depends on the energy bins that are used and should account for energy dispersion as well as resolution. In this step the energy channels are fitted simultaneously, as it is impossible to incorporate energy dispersion in a way that fits every energy channel individually. This fitting can be done in different spectral fitting software like 3ML or XSPEC. In this work we use 3ML.

\subsection{Synchrotron Model}
\label{synch}
We use pynchrotron (cite github) to calculate the spectrum from a physical, time dependent spectral synchrotron model. This model was previously used to successfully fit many single pulse GRBs with GBM in cite syn paper, which also includes a detailed description of the model. Here we only want to summarize the main points shortly.

The core of the model is the assumption of some generic acceleration mechanism, that constantly produces an electron spectrum with a power law shape $N(\gamma )\propto \gamma^{-p}$ between $\gamma_{inj}$ and $\gamma_{max}$. These electrons are cooled via synchrotron cooling in a magnetic field and produce radiation in this process. The resulting photon spectrum is the sum of the photon spectra at every time step in the cooling process and is defined by five quantities:

@@

There exists a strong degeneracy between $B$ and $\gamma_{inj}$, as their combination sets the peak of the photon spectrum. Therefore one usually fixes $\gamma_{inj}=10^{5}$ and only fits for $B$. It is important to keep in mind, that the results for $\gamma_{max}$ and $\gamma_{cool}$ are determined relative to the fixed $\gamma_{inj}$.

\subsection{Model Checking}
\label{PPC}
To check if the fit of a model is a good description of the measured data is a very complicated topic and highly debated in the statistics community. Measurements like reduced $\chi^{2}$ use many assumptions, like for example that all the probability distributions in the problem are described as normal distributions, which is not the case here, as the measurement process is a Poisson process. Therefore we decided to use Posterior Predictive Checks (PPC) and Quantile-Quantile (QQ) plots. For the PPCs one simulate new data from the full posterior of the fit as well as the measurement process (see Eq. \ref{eq:ppc}) and compare them to the observed data. QQ plots use the same simualtion process, but instead of comparing the observed data of every energy channel individually to the simulated data, one compares the cumulative counts of the observation and the simulations over energy channels. (cite) QQ plots are very sensitive to weak systematic deviations of the model from the data.

@@

We analyze the data from the two instruments independently first with the empirical Band function, to check if the results for SPI and GBM are in agreement as well as show a comparison between the fit results with PySPI and the official software tools for GRB analysis within OSA. Then we fit the SPI and GBM data simultaneously with a physical synchrotron model, and show that it can reduce the allowed parameter space significantly, compared to the individual fits. For all GBM and PySPI fits we added effective area correction parameters in the fit, allowing the total effective area of the individual detectors to vary a bit with respect to each other, to account for slightly different calibrations. To do this, we always fix the response of one of the detectors to the normal response of that detectors and fit one parameter for every other detector, which gets multiplied with the response in the fit. We constrain the effective area correction parameters to be between 0.7 and 1.3.

All the fits in this section are done with 3ML (cite) and the Bayesian sampling algorithm MultiNest (cite) to create full posterior distributions of the parameters. To visualize the posterior distribution we use ChainConsumer (cite) with the output from MultiNest to create corner plots.

@@

\begin{figure}
  \begin{centering}
    \includegraphics[width=1\linewidth]{figures/pyspi_vs_osa_band.pdf}
    \caption{Corner plot for the SPI fit with PySPI and the official tools. The spectral model in this fit is a Band function. It shows, that the results are in agreement, but PySPI can constrain the parameters better.}
    \label{fig:corner_osa_pyspi_band}
  \end{centering}
\end{figure}

@@

We analyzed the data of GRB120711A with the data from INTGRAL/SPI and Fermi/GBM. The results for SPI with PySPI are in agreement with those from the official analysis software, but more constraining. Also the resulting spectral shape for the fits with GBM and SPI data are in agreement and the relative calibration difference between SPI and GBM we conclude from the fits, is $\approx$ 10\% or less. We show that the GBM as well as the SPI data for GRB120711A can be fitted well with a physical, time dependent synchrotron model and that combining the data in a simultaneous fit improves the parameter constrains.

Generally PySPI can improve the scientific output from SPI for GRBs significantly. Due to a new analysis method, that folds the source model directly into the data space of every SPI detector with the full response of the individual detectors, the results are more constraining than the results with the official analysis software.
